Computing Exact Restricted Isometry Constants via\\ Mixed Integer Semidefinite Programming

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
introduced short form "SDP"

\noindent
\begin{abstract}
  One of the fundamental tasks in compressed sensing is finding the
  sparsest solution to an underdetermined system of linear equations. It is
  well known that although this problem is \NP-hard, under certain
  conditions it can be solved by using a linear program which minimizes the
  1-norm. The restricted isometry property has been one of the key
  conditions in this context. However, computing the best constants for
  this condition is itself \NP-hard. In this paper we propose a
  mixed-integer semidefinite programming approach for computing these
  optimal constants. This also subsumes sparse principal component
  analysis. Computational results with this approach allow to evaluate
  earlier SDP relaxations and show that their quality is limited.
\end{abstract}

@@

This motivated d'Aspremont et al.~\cite{Asp08,Asp07} to propose several
semidefinite relaxations to compute upper bounds on the restricted isometry
constants; small upper bounds suffice to guarantee
$\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
there exist cases in which the bounds are too weak to give a guarantee.

@@

As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
\begin{align}\label{Asp08}
 \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1, \tag{A2-Primal}\\
 & X \succeq 0,  \nonumber \\
 & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n, \nonumber
\end{align}
for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
\begin{align}\label{Asp08Dual}
  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \nonumber \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \tag{A2-Dual} \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n, \nonumber
\end{align}
which has only half as many variables as \eqref{Asp08} and one fewer linear and semidefinite constraint. Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for \eqref{phi} will always be $x=0$.

@@

\noindent
In this section we propose a MISDP formulation to compute the optimal
constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
\eqref{upperRIP}, respectively. We will again start with the non-convex quadratic problem~\eqref{QP}.
For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in \eqref{QP} can equivalently be written using
binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i
\leq k$. We can thus equivalently rewrite \eqref{QP} as
\begin{align}\label{Rk1MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
 & \sum_{i=1}^n z_i \leq k,\tag{Rk1MISDP}\\
 & \Rk(X) = 1, \nonumber \\
 & X \succeq 0,  \nonumber \\
 & z \in \{0,1\}^n. \nonumber
\end{align}
We then relax the non-convex rank constraint to arrive at the following
MISDP formulation:
\begin{align}\label{MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
 & \sum_{i=1}^n z_i \leq k, \tag{MISDP} \\
 & X \succeq 0,  \nonumber \\
 & z \in \{0,1\}^n. \nonumber
\end{align}
In the following, we will show that \eqref{MISDP} actually produces the
exact value of $\alpha_k$. We will use the following result about the
relationship between the rank of solutions of semidefinite programs and the
dimension of corresponding faces of the feasible set~$P$, i.e., subsets $F
\subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply
$y,z \in F$:

>>>>

\noindent
\begin{abstract}
  One of the fundamental tasks in compressed sensing is finding the
  sparsest solution to an underdetermined system of linear equations. It is
  well known that although this problem is \NP-hard, under certain
  conditions it can be solved by using a linear program which minimizes the
  1-norm. The restricted isometry property has been one of the key
  conditions in this context. However, computing the best constants for
  this condition is itself \NP-hard. In this paper we propose a
  mixed-integer semidefinite programming approach for computing these
  optimal constants. This also subsumes sparse principal component
  analysis. Computational results with this approach allow to evaluate
  earlier semidefinite relaxations and show that their quality is limited.
\end{abstract}

@@

This motivated d'Aspremont et al.~\cite{Asp08,Asp07} to propose several
relaxations based on semidefinite programming (SDP) to compute upper bounds on the restricted isometry
constants; small upper bounds suffice to guarantee
$\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
there exist cases in which the bounds are too weak to give a guarantee.

@@

As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
\begin{align}\label{Asp08}
 \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1, \tag{A2-Primal}\\
 & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n, \nonumber
\end{align}
for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
\begin{align}\label{Asp08Dual}
  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \nonumber \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \tag{A2-Dual} \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n, \nonumber
\end{align}
which has only half as many variables as \eqref{Asp08} and one fewer linear and semidefinite constraint. Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for \eqref{phi} will always be $x=0$.

@@

\noindent
In this section we propose a MISDP formulation to compute the optimal
constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
\eqref{upperRIP}, respectively. We will again start with the non-convex quadratic problem~\eqref{QP}.
For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in \eqref{QP} can equivalently be written using
binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i
\leq k$. We can thus equivalently rewrite \eqref{QP} as
\begin{align}\label{Rk1MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
 & \sum_{i=1}^n z_i \leq k,\\
 & \Rk(X) = 1, \nonumber \\
 & X \succeq 0,  \nonumber \\
 & z \in \{0,1\}^n. \nonumber
\end{align}
We then relax the non-convex rank constraint to arrive at the following
MISDP formulation:
\begin{align}\label{MISDP}
 \min \quad & \Tr(A\T A X) \nonumber \\
 \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
 & \sum_{i=1}^n z_i \leq k, \tag{MISDP} \\
 & X \succeq 0,  \nonumber \\
 & z \in \{0,1\}^n. \nonumber
\end{align}
In the following, we will show that \eqref{MISDP} actually produces the
exact value of $\alpha_k$. We will use the following result about the
relationship between the rank of solutions of semidefinite programs and the
dimension of corresponding faces of the feasible set~$P$, i.e., subsets $F
\subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply
$y,z \in F$:
